---
title: "Homework9"
author: "Noah W.K. Mattheis"
date: "2025-03-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 

First step, load up proper libraries and packages

```{r}
# loading up libraries
library(devtools)
install_github("ngotelli/upscaler")
library(upscaler)
```

Starting with pseudo-code, generate functions for 1) Cleaning the data for any empty/missing cases, 2) Extract the year from each file name, 3) Calculate Abundance for each year (Total number of individuals found), 4) Calculate Species Richness for each year(Number of unique species found), 5) Run a simple regression model for Species Richness (S) vs. Abundance for every year, 6) Generate histograms for both Abundance and Species Richness (S) and store the plots


Creating Functions for For Loop for reading in data
```{r}
add_folder("Functions")
build_function(c("clean_data",
                 "get_abundance",
                 "get_year",
                 "species_richness",
                 "simple_regression",
                 "gen_hist"))

# When calling back in functions, make sure to look into functions folder to see the name given to these functions
source_batch("Functions")

# Running each function template
clean_data()
gen_hist()
get_abundance()
get_year()
simple_regression()
species_richness()

# Testing all the functions
temp_extract <- get_year("C:/Users/noahw/OneDrive/Desktop/UVM/Spring 2025/BIOL 6100/R Code/MattheisBIOL6100/NEON_count-landbird (1)/NEON_count-landbird/NEON.D01.BART.DP1.10003.001.2015-06.basic.20250129T000730Z.RELEASE-2025", "*.brd_countdata.*")

temp_extract
placeholder <- read.csv(temp_extract)
q <- clean_data(placeholder)
class(placeholder)

placeholder2 <- read.csv(temp_extract)
species <- species_richness(placeholder2)
abundance <- get_abundance(placeholder2)
simple_regression(placeholder2, species, abundance)
gen_hist(placeholder2, placeholder2$uid, placeholder2$scientificName)

paste0("C:/Users/noahw/OneDrive/Desktop/UVM/Spring 2025/BIOL 6100/R Code/MattheisBIOL6100/NEON_count-landbird (1)/NEON_count-landbird/NEON.D01.BART.DP1.10003.001.20",15, "-06.basic.20250129T000730Z.RELEASE-2025")

```

Now, we uploaded some data from Neonscience.org and are working with the BART from years 2013 to 2023. Now is time to merge the countbird files for each year into one file together and use our pseudo-code and functions to help

```{r}

file_list <- list.files(path ="C:/Users/noahw/OneDrive/Desktop/UVM/Spring 2025/BIOL 6100/R Code/MattheisBIOL6100/NEON_count-landbird (1)/NEON_count-landbird" ) # set working directory to inside data file so you do not have to specify path
file_list[2]

temp_list <- list()
new_data <- data_frame(
  "year_name" = character(),
  "spec_num" = numeric(),
  "abundance_num" = numeric()
)



for (i in 1:length(file_list)) {
  temp_list[i] <- get_year(paste0("C:/Users/noahw/OneDrive/Desktop/UVM/Spring 2025/BIOL 6100/R Code/MattheisBIOL6100/NEON_count-landbird (1)/NEON_count-landbird/NEON.D01.BART.DP1.10003.001.20",i + 14, "-06.basic.20250129T000730Z.RELEASE-2025"), ".brd_countdata.")
  temp_list_name <- temp_list[i]
  temp_list_name 
  placeholder <- read.csv(unlist(temp_list[i]))
  clean_placeholder <- clean_data(placeholder)
  species <- species_richness(clean_placeholder)
  abundance <- get_abundance(clean_placeholder)
  new_data[i, "year_name"] <-  temp_list_name 
  new_data[i, "spec_num"] <- abundance
  new_data[i, "abundance_num"] <- species
  
  
}

head(new_data)

# Need to extract p-value, adj r squared
# look up how to pull out these values from a lm
# then put them into a data frame

simple_reg <- (simple_regression(new_data, 
                                      new_data$spec_num, 
                                      new_data$abundance_num))
regression_data <- data.frame()


simple_reg$coefficients[8]
simple_reg$r.squared
simple_reg$adj.r.squared

regression_data <- c("p-value" = simple_reg$coefficients[8],
"r squared" = simple_reg$r.squared,
"adj r squared" = simple_reg$adj.r.squared)

write.csv(regression_data, file = "regression_data")


gen_hist(new_data, new_data$abundance_num, new_data$spec_num)

```
